\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
changed a single word

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\subsection{Boolean circuits}
We consider the problem of learning
to build  Boolean circuits out of logic gates.
If we gave our agent the full repertoire of
circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
Instead, we gave \system only a NAND gate,
and tasked it with building a set of 250 random circuits made out of
AND, NOT, and OR gates, along with a multiplexer.
\system observes the input/output truth table of the circuits,
and we define the likelihood 
$  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]$,
where $x$ is a truth table and $p$ a program implementing a circuit.

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string methods provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot programs for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \code{0}, \code{+1},\code{-1},     \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  \\
    \code{split}, \code{join}, \code{map}, 
    \code{+}, \code{chr->str}, \code{trim}, \code{nth}\\
    \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\code{c}$) \,=\, $\code{(+ (chr->str c))}\\
      $f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ (nth i (split c s))}\\
        \phantom{$f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ }}\code{(chr->str c))}\\
      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
        \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
        \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\code{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some learned DSL primitives.}\label{textPrimitives}
  \end{table}

\begin{figure}
  \centering\begin{tabular}{ll|ll}\toprule
    Input&Output&Input&Output\\\midrule
cqbHT$>$aIKd$<$W& 	AIKD&BgZQ..PBBnD & BgZQ. \\
YBW$>$TIg$<$WS 	&TIG& qqfgg..mNIx & qqfgg. 
%DkI$>$DwKI$<$xaPoc& 	DWKI\\
%KfvTh$>$Dbl$<$OIJso 	&DBL
\\\midrule
\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split '>' } }&
\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s '.' 0)}}
\\
\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) '<')}}
\\\bottomrule
  \end{tabular}

@@

\subsection{List functions}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 225 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 225 tasks.

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$,
e.g. did the agent correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
often dramatically (string editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports our view of the recognition model as a way of accelerating the search over programs.

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\subsection{Boolean Circuits}
We consider the problem of learning
to build  Boolean circuits out of logic gates.
If we gave our agent the full repertoire of
circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
Instead, we gave \system only a NAND gate,
and tasked it with building a set of 250 random circuits made out of
AND, NOT, and OR gates, along with a multiplexer.
\system observes the input/output truth table of the circuits,
and we define the likelihood 
$  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]$,
where $x$ is a truth table and $p$ a program implementing a circuit.

@@

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string methods provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot programs for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \code{0}, \code{+1},\code{-1},\codechar{},\codechar{,},\codechar{ },\codechar{<},\codechar{>}  \\
    \code{split}, \code{join}, \code{map}, 
    \code{+}, \code{chr->str}, \code{trim}, \code{nth}\\
    \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\code{c}$) \,=\, $\code{(+ (chr->str c))}\\
      $f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ (nth i (split c s))}\\
        \phantom{$f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ }}\code{(chr->str c))}\\
      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (+1 0) c s))} \\
        \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (+1 0) (split c (nth 0 s)))}\\
        \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\code{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some learned DSL primitives.}\label{textPrimitives}
  \end{table}

\begin{figure}
  \centering\begin{tabular}{ll|ll}\toprule
    Input&Output&Input&Output\\\midrule
cqbHT$>$aIKd$<$W& 	AIKD&BgZQ..PBBnD & BgZQ. \\
YBW$>$TIg$<$WS 	&TIG& qqfgg..mNIx & qqfgg. 
%DkI$>$DwKI$<$xaPoc& 	DWKI\\
%KfvTh$>$Dbl$<$OIJso 	&DBL
\\\midrule
\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split }\codechar{>}\code{ } }&
\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s }\codechar{.}\code{ 0)}}
\\
\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) }\codechar{<}\code{)}}
\\\bottomrule
  \end{tabular}

@@

\subsection{List Functions}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 225 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 225 tasks.

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$,
e.g. did the agent correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (string editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports our view of the recognition model as a way of accelerating the search over programs.
